---
title: "Football Match Network Analysis"
author: "Umut Evren"
date: "2024-06-22"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: united
    highlight: tango
    code_folding: show
---

```{r setup, include=FALSE}
# Load required libraries
library(igraph)        # For network analysis and visualization
library(visNetwork)    # For interactive network visualization
library(dplyr)         # For data manipulation
library(ggraph)        # For network visualization
library(tidygraph)     # For network data manipulation
library(gt)            # For creating beautiful tables
library(circlize)      # For chord diagrams

# Set global options
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 7
)
```

# Introduction

This document presents a comprehensive network analysis of football match data, focusing on passing patterns and player interactions between two teams. The analysis includes various network metrics, visualizations, and statistical measures to understand team dynamics and player roles.

# Data Loading and Preparation

```{r data_loading}
# Load team matrices
df_team_a <- read.csv(
  file = "/Users/umutevren/Downloads/teama_matrix.csv", 
  row.names = 1, 
  check.names = FALSE, 
  stringsAsFactors = FALSE
)

df_team_b <- read.csv(
  file = "/Users/umutevren/Downloads/teamb_matrix.csv", 
  row.names = 1, 
  check.names = FALSE, 
  stringsAsFactors = FALSE
)

# Display data structure
str(df_team_a)
str(df_team_b)
```

# Network Construction

## Data Preparation

```{r network_preparation}
# Extract player columns and create matrices
player_columns_a <- grep("^\\d+$", colnames(df_team_a), value = TRUE)
player_columns_b <- grep("^\\d+$", colnames(df_team_b), value = TRUE)

matrix_team_a <- df_team_a %>%
  select(all_of(player_columns_a)) %>%
  as.matrix()

matrix_team_b <- df_team_b %>%
  select(all_of(player_columns_b)) %>%
  as.matrix()

# Create network objects
net_adj_a <- graph.adjacency(matrix_team_a, weighted = TRUE, diag = FALSE)
net_adj_b <- graph.adjacency(matrix_team_b, weighted = TRUE, diag = FALSE)
```

## Network Properties

```{r network_properties}
# Set vertex labels
V(net_adj_a)$label <- as.character(df_team_a$Player)
V(net_adj_b)$label <- as.character(df_team_b$Player)

# Set vertex colors
V(net_adj_a)$color <- c("blue","yellow","blue","red","red","blue","green","red","yellow","yellow","red")
V(net_adj_b)$color <- c("green","yellow","red","blue","yellow","yellow","yellow","red","red","blue","blue")

# Set vertex sizes based on degree
V(net_adj_a)$size <- degree(net_adj_a)
V(net_adj_b)$size <- degree(net_adj_b)

# Set edge properties
E(net_adj_a)$width <- E(net_adj_a)$weight
E(net_adj_b)$width <- E(net_adj_b)$weight

# Highlight strong ties
Amed <- median(E(net_adj_a)$weight)
Bmed <- median(E(net_adj_b)$weight)
E(net_adj_a)$color <- ifelse(E(net_adj_a)$weight > Amed, "gray30", "gray70")
E(net_adj_b)$color <- ifelse(E(net_adj_b)$weight > Bmed, "gray30", "gray70")
```

# Network Visualization

## Basic Network Plots

```{r basic_network_plots}
# Save combined network plot
png("/Users/umutevren/Documents/v001_team_networks.png", width = 800, height = 600)
set.seed(1234)
par(mfrow = c(1,2))
plot(net_adj_a, main = "Go Ahead Team A")
plot(net_adj_b, main = "Go Ahead Team B")
par(mfrow = c(1,1))
dev.off()
```

## Custom Layout Network Plots

```{r custom_layout}
# Define custom coordinates for better visualization
A_coords <- data.frame(
  row.names = V(net_adj_a)$label,
  "x" = c(-1, 0.2, -0.75, 0, 0.75, 0.5, 0, 0, -0.5, -0.5, -0.75),
  "y" = c(0.50, 0.2, 0.8, -0.5, -0.5, 1, -1, -0.3, 0, 0.45, -0.5)
)

B_coords <- data.frame(
  row.names = V(net_adj_b)$label,
  "x" = c(0, 0.5, -0.74, 0.78, 0, 0.15, -0.5, 0, 0.75, -0.73, 0),
  "y" = c(-1, 0, -0.25, 0.8, 0, 0.3, 0.25, -0.5, -0.25, 0.79, 0.75)
)

# Plot networks with custom layout
plot(net_adj_a,
     main = "Team A",
     layout = as.matrix(A_coords),
     vertex.label.dist = 2,
     edge.arrow.size = 0.4,
     edge.curved = 0.1,
     vertex.size = 10)

plot(net_adj_b,
     main = "Team B",
     layout = as.matrix(B_coords),
     vertex.label.dist = 2,
     edge.arrow.size = 0.4,
     edge.curved = 0.1,
     vertex.size = 10)
```

## Advanced Network Visualizations

```{r advanced_visualizations}
# Create interactive network visualization for Team B
nodes_b_v2 <- data.frame(
  id = V(net_adj_b)$name,
  label = V(net_adj_b)$name,
  color = V(net_adj_b)$color
)

edges_b_v2 <- data.frame(
  from = as.character(ends(net_adj_b, E(net_adj_b))[,1]),
  to = as.character(ends(net_adj_b, E(net_adj_b))[,2]),
  arrows = "to"
)

# Interactive network plot
visNetwork(nodes_b_v2, edges_b_v2, main = "Team B") %>%
  visNodes(size = 30) %>%
  visEdges(arrows = 'to') %>%
  visOptions(highlightNearest = TRUE) %>%
  visInteraction(navigationButtons = TRUE)
```

# Network Analysis

## Betweenness Centrality

```{r betweenness_analysis}
# Calculate betweenness centrality
bt_adj_a <- graph.adjacency(matrix_team_a, weighted = TRUE, diag = FALSE)
bt_adj_b <- graph.adjacency(matrix_team_b, weighted = TRUE, diag = FALSE)

bt_score_a <- betweenness(bt_adj_a)
bt_score_b <- betweenness(bt_adj_b)

# Create data frames for visualization
bt_df_a <- data.frame(Node = node_names_a, Betweenness = bt_score_a)
bt_df_b <- data.frame(Node = node_names_b, Betweenness = bt_score_b)

# Prepare data for tables
bt_df_a_v2 <- bt_df_a %>%
  arrange(desc(Betweenness)) %>%
  mutate(color = "")

bt_df_b_v2 <- bt_df_b %>%
  arrange(desc(Betweenness)) %>%
  mutate(color = "")

# Create formatted tables
bt_a_table <- gt(data = bt_df_a_v2) %>%
  data_color(
    columns = vars(Betweenness),
    target_columns = vars(color),
    method = "numeric",
    palette = "viridis",
    domain = range(bt_df_a_v2$Betweenness)
  ) %>%
  cols_label(
    Node = "Node",
    Betweenness = "Betweenness",
    color = ""
  ) %>%
  opt_vertical_padding(scale = 0.65) %>%
  tab_caption(caption = "Team A (Betweenness)")

bt_b_table <- gt(data = bt_df_b_v2) %>%
  data_color(
    columns = vars(Betweenness),
    target_columns = vars(color),
    method = "numeric",
    palette = "viridis",
    domain = range(bt_df_b_v2$Betweenness)
  ) %>%
  cols_label(
    Node = "Node",
    Betweenness = "Betweenness",
    color = ""
  ) %>%
  opt_vertical_padding(scale = 0.65) %>%
  tab_caption(caption = "Team B (Betweenness)")

# Display tables
bt_a_table
bt_b_table

# Save tables
gtsave(bt_b_table, filename = "bt_b_table.png")
gtsave(bt_a_table, filename = "bt_a_table.png")
```

## PageRank Analysis

```{r pagerank_analysis}
# Calculate PageRank
page_rank_result_a <- page_rank(net_adj_a, directed = TRUE, damping = 0.85)
page_rank_result_b <- page_rank(net_adj_b, directed = TRUE, damping = 0.85)

# Create data frames
page_rank_df_a <- data.frame(
  Node = node_names_a,
  PageRank = page_rank_result_a$vector,
  Group = "Team A"
)

page_rank_df_b <- data.frame(
  Node = node_names_b,
  PageRank = page_rank_result_b$vector,
  Group = "Team B"
)

# Prepare data for visualization
page_rank_df_a_v2 <- page_rank_df_a %>%
  arrange(desc(PageRank)) %>%
  mutate(color = "")

page_rank_df_b_v2 <- page_rank_df_b %>%
  arrange(desc(PageRank)) %>%
  mutate(color = "")

# Create formatted tables
pr_a_table <- gt(data = page_rank_df_a_v2) %>%
  cols_hide(columns = vars(Group)) %>%
  data_color(
    columns = vars(PageRank),
    target_columns = vars(color),
    method = "numeric",
    palette = "viridis",
    domain = range(page_rank_df_a_v2$PageRank)
  ) %>%
  cols_label(
    Node = "Node",
    PageRank = "PageRank",
    color = ""
  ) %>%
  opt_vertical_padding(scale = 0.65) %>%
  tab_caption(caption = "Team A (PageRank)")

pr_b_table <- gt(data = page_rank_df_b_v2) %>%
  cols_hide(columns = vars(Group)) %>%
  data_color(
    columns = vars(PageRank),
    target_columns = vars(color),
    method = "numeric",
    palette = "viridis",
    domain = range(page_rank_df_b_v2$PageRank)
  ) %>%
  cols_label(
    Node = "Node",
    PageRank = "PageRank",
    color = ""
  ) %>%
  opt_vertical_padding(scale = 0.65) %>%
  tab_caption(caption = "Team B (PageRank)")

# Display tables
pr_a_table
pr_b_table

# Save tables
gtsave(pr_b_table, filename = "pr_b_table.png")
gtsave(pr_a_table, filename = "pr_a_table.png")
```

# Chord Diagrams

## Data Preparation for Chord Diagrams

```{r chord_preparation}
# Prepare Team A data
df_team_a_v2 <- df_team_a[, -c(ncol(df_team_a) - 1, ncol(df_team_a))]
rownames(df_team_a_v2) <- df_team_a_v2[, 1]
df_team_a_v2 <- df_team_a_v2[, -1]

matrix_df_team_a_v2 <- as.matrix(df_team_a_v2)
colnames(matrix_df_team_a_v2) <- rownames(matrix_df_team_a_v2)

pairs_matrix_df_team_a_v2 <- as.data.frame(as.table(matrix_df_team_a_v2)) %>%
  filter(Freq > 0) %>%
  rename(from = Var1, to = Var2, value = Freq)

# Prepare Team B data
df_team_b_v2 <- df_team_b[, -c(ncol(df_team_b) - 1, ncol(df_team_b))]
rownames(df_team_b_v2) <- df_team_b_v2[, 1]
df_team_b_v2 <- df_team_b_v2[, -1]

matrix_df_team_b_v2 <- as.matrix(df_team_b_v2)
colnames(matrix_df_team_b_v2) <- rownames(matrix_df_team_b_v2)

pairs_matrix_df_team_b_v2 <- as.data.frame(as.table(matrix_df_team_b_v2)) %>%
  filter(Freq > 0) %>%
  rename(from = Var1, to = Var2, value = Freq)
```

## Create Chord Diagrams

```{r chord_diagrams}
# Team A Chord Diagram
scaled_lwd_a_2 <- pairs_matrix_df_team_a_v2$value / max(pairs_matrix_df_team_a_v2$value) * 10

png("team_a_chord_v8.png", width = 1600, height = 1600, res = 150)
circos.clear()
chordDiagram(pairs_matrix_df_team_a_v2,
             transparency = 0.5,
             annotationTrack = "grid",
             preAllocateTracks = 1,
             link.lwd = scaled_lwd_a_2,
             annotationTrackHeight = mm_h(4),
             link.border = 0.1)

circos.track(track.index = 1, panel.fun = function(x, y) {
  circos.text(CELL_META$xcenter, CELL_META$ylim[1], CELL_META$sector.index,
              facing = "clockwise", niceFacing = TRUE, adj = c(0, 0.5),
              cex = 1.2, col = "black")
}, bg.border = NA)
dev.off()

# Team B Chord Diagram
scaled_lwd_b_2 <- pairs_matrix_df_team_b_v2$value / max(pairs_matrix_df_team_b_v2$value) * 10

png("team_b_chord_v2.png", width = 1600, height = 1600, res = 150)
circos.clear()
chordDiagram(pairs_matrix_df_team_b_v2,
             transparency = 0.5,
             annotationTrack = "grid",
             preAllocateTracks = 1,
             link.lwd = scaled_lwd_b_2,
             annotationTrackHeight = mm_h(4),
             link.border = 0.1)

title("Team B - Pass Pairs", cex.main = 1, line = -0.05)

circos.track(track.index = 1, panel.fun = function(x, y) {
  circos.text(CELL_META$xcenter, CELL_META$ylim[1], CELL_META$sector.index,
              facing = "clockwise", niceFacing = TRUE, adj = c(0, 0.5),
              cex = 0.8, col = "black")
}, bg.border = NA)
dev.off()
```

# Advanced Network Visualization

```{r advanced_network}
# Create advanced network visualization for Team B
ggraph(net_adj_b, layout = B_coords) +
  geom_edge_fan(aes(width = width, colour = color),
                alpha = 0.6,
                arrow = arrow(length = unit(4, 'mm')),
                start_cap = circle(3, 'mm'),
                end_cap = circle(3, 'mm')) +
  geom_node_point(aes(label = label, size = size, colour = color)) +
  geom_node_text(aes(label = label), repel = TRUE) +
  scale_edge_color_manual(values = c("gray30", "gray70")) +
  ggtitle("Team B") +
  theme_graph() +
  theme(legend.position = "none")

ggsave("old_school_Team_B_graph.png", width = 10, height = 7)
```

# Conclusion

This analysis provides a comprehensive view of team dynamics through network analysis. The visualizations and metrics help understand:

1. Player interactions and passing patterns
2. Key players based on betweenness centrality
3. Player importance through PageRank analysis
4. Team structure and flow through chord diagrams
5. Network properties and characteristics

The results can be used to:
- Identify key players and their roles
- Understand team strategies and patterns
- Compare team structures and dynamics
- Make data-driven decisions for team improvement 